### 2022-03-22 docs/ms.tex
add, with professional easter egg

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
\author{true}
\date{}

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Steven V. Miller (Clemson University)},
             pdfkeywords = {software, statistical analysis, peace science, reproducibility},
            pdftitle={\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
   \MakeUppercase{Steven V. Miller}, \small{Clemson University}

@@

\texttt{\{peacesciencer\}} is designed to address these problems. Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers. Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch. Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. \texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts. All are done with the maximum possible transparency. The project is available for public view on Github (\url{https://github.com/svmiller/peacesciencer/}). The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created. The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

\begin{longtable}[t]{>{}lc>{\raggedright\arraybackslash}p{20em}}
\caption{\label{tab:tab-addfunctions}Functions that Add Data}\\
\toprule
\textbf{Function} & \textbf{Type} & \textbf{Description}\\
\midrule
\ttfamily{add\_archigos()} & D,S & Add Archigos political leader information to a data frame\\
\ttfamily{add\_atop\_alliance()} & D, LD & Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame\\
\ttfamily{add\_capital\_distance()} & D, L, LD, S & Add capital-to-capital distance to a data frame\\
\ttfamily{add\_ccode\_to\_gw()} & D, L, LD, S & Match CoW state codes to G-W state codes\\
\ttfamily{add\_contiguity()} & D, L, LD, S & Add CoW direct contiguity information to a data frame\\
\ttfamily{add\_cow\_alliance()} & D, LD & Add CoW alliance data to a data frame\\
\ttfamily{add\_cow\_majors()} & D, L, LD, S & Add CoW major power information to a data frame\\
\ttfamily{add\_cow\_mids()} & D & Add CoW-MID data to a dyad-year data frame\\
\ttfamily{add\_cow\_trade()} & D, L, LD, S & Add CoW trade data to a data frame\\
\ttfamily{add\_cow\_wars()} & D, S & Add CoW war data to a dyad-year or state-year data frame\\
\ttfamily{add\_creg\_fractionalization()} & D, L, LD, S & Add fractionalization/polarization estimates from CREG to a data frame\\
\ttfamily{add\_democracy()} & D, L, LD, S & Add democracy information to a data frame\\
\ttfamily{add\_fpsim()} & D, LD & Add dyadic foreign policy similarity measures to a data frame\\
\ttfamily{add\_gml\_mids()} & D, L, LD, S & Add GML MID data to a data frame\\
\ttfamily{add\_gwcode\_to\_cow()} & D, L, LD, S & Match G-W state codes to CoW state codes\\
\ttfamily{add\_igos()} & D, S & Add CoW IGO data to a data frame\\
\ttfamily{add\_lead()} & L, LD & Add leader experience/attributes (LEAD) to data\\
\ttfamily{add\_lwuf()} & L, LD & Add leader willingness to use force estimates to data\\
\ttfamily{add\_minimum\_distance()} & D, L, LD, S & Add minimum distance data to data frame\\
\ttfamily{add\_nmc()} & D, L, LD, S & Add Correlates of War National Military Capabilities data\\
\ttfamily{add\_peace\_years()} & D, S & Add 'peace years' to dyad-year/state-year conflict data\\
\ttfamily{add\_rugged\_terrain()} & D, L, LD, S & Add rugged terrain information to a data frame\\
\ttfamily{add\_sdp\_gdp()} & D, L, LD, S & Add (surplus, gross) domestic product data to a data frame\\
\ttfamily{add\_spells()} & D, L, LD, S & Add duration 'spells' to a data frame\\
\ttfamily{add\_strategic\_rivalries()} & D, S & Add strategic rivalry information to a data frame\\
\ttfamily{add\_ucdp\_acd()} & S & Add UCDP Armed Conflict Data to state-year data frame\\
\ttfamily{add\_ucdp\_onsets} & S & Add UCDP onsets to state-year data\\
\ttfamily{filter\_prd()} & D, LD & Filter dyad-year data to just politically relevant dyads\\
\ttfamily{wc\_duration()} & C & Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration\\
\ttfamily{wc\_fatality()} & C & Whittle Duplicate Conflict-Years by Highest Fatality\\
\ttfamily{wc\_hostility()} & C & Whittle Duplicate Conflict-Years by Highest Hostility\\
\ttfamily{wc\_jds()} & C & Whittle Duplicate Conflict-Years by Just Dropping Something\\
\ttfamily{wc\_onsets()} & C & Whittle Uniqute Conflict Onset-Years from Conflict-Year Data\\
\ttfamily{wc\_recip()} & C & Whittle Duplicate Conflict-Years by Conflict Reciprocation\\
\ttfamily{wc\_stmon()} & C & Whittle Duplicate Conflict-Years by Lowest Start Month\\
\bottomrule
\end{longtable}
\endgroup{}

@@

\texttt{\{peacesciencer\}} is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2. \texttt{\{peacesciencer\}} is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

\texttt{\{peacesciencer\}}'s functions work out of the box, though users should find their experience augmented by two additional downloads. First, RStudio offers an IDE that serves as a user-friendly graphical user interface (GUI) over what is, at its core, a programming language with a command-line interface. RStudio's design will make it much easier for users to experiment with \texttt{\{peacesciencer\}}'s functionality and read its documentation to assist them with the use of these functions. The second additional download is \texttt{\{tidyverse\}}, itself a suite of packages that share a common form and design (Wickham et al., 2019). \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}}, and \texttt{\{peacesciencer\}} can work without it, but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality. A user can open an R session by way of RStudio and install both packages as follows.\footnote{R users will recognize that pound signs (\texttt{\#}) allow for commenting code. Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing. In subsequent code snippets, \texttt{\#\textgreater{}} will indicate console output the user can expect from these functions.}

@@

A user who approaches \texttt{\{peacesciencer\}} with a project in mind will see they can better tailor this function to what they want. Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data. They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward. Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).

@@

Researchers using \texttt{\{peacesciencer\}} to create data for their research project should start with one of these ``create'' functions. Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.

@@

Researchers should see that \texttt{\{peacesciencer\}}'s functionality can scale up nicely from there. For example, the following would round out the kind of information necessary to replicate Bremer's (1992) famous ``dangerous dyads'' analysis by adding information about national material capabilities for both states in the dyad (\texttt{add\_nmc()}), estimates of democracy for both states in the dyad (\texttt{add\_democracy()}), information about alliance commitments in the dyad-year (\texttt{add\_cow\_alliance()}, by way of Gibler (2009)), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by Anders, Fariss \& Markowitz (2020). Whereas \texttt{add\_sdp\_gdp()} is the last command in the pipe-based workflow, the \texttt{\{peacesciencer\}} call ends by assigning to an object called \texttt{Data}. This type of assignment is done with the ``right hand'' assignment operator (i.e.~\texttt{-\textgreater{}}).

@@

If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like \texttt{haven::write\_dta(Data,\ "my-data.dta")} and import it into Stata when they are done. No matter, \texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the capabilities data. In this particular application, it has already created the main data required for a replication of Bremer (1992). There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad. Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators. Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but \texttt{\{peacesciencer\}} greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

@@

\texttt{\{peacesciencer\}} is capable of generating data for replications of analyses at multiple levels beyond dyad-year. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{},\ subset\_years\ =\ c(1946:2019))}, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in \texttt{\{peacesciencer\}} include just the observations in that time frame). Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states. \texttt{add\_ucdp\_acd()} takes three arguments: \texttt{type}, \texttt{issue}, and \texttt{only\_wars}. \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information. Options include ``extrasystemic'', ``interstate'', ``intrastate'', and ``II'' (short for ``internationalized intrastate''). If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}). \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include ``territory'', ``government'', and ``both''. If no \texttt{issue} is specified, the function returns information for all conflicts regardless of the particular issue. \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}. The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE))} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}

@@

Finally, we can add some covariates of interest to these data. \texttt{add\_spells()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}. \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores). \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. \texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss \& Markowitz (2020) simulations. Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by Nunn \& Puga (2012) and the second is the Gibler \& Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon \& Laitin (2003)). At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.

@@

\caption{\label{tab:tab-cw}A Civil Conflict Analysis of Gleditsch-Ward State-Years in \{peacesciencer\}}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8cm}cc}
\toprule
\textbf{ } & \textbf{All UCDP Conflicts} & \textbf{Wars Only}\\
\midrule
GDP per Capita (Lagged) & \num{-0.285}* & \num{-0.343}*\\
 & (\num{0.110}) & (\num{0.172})\\
Population Size (Lagged) & \num{0.229}* & \num{0.272}*\\
 & (\num{0.067}) & (\num{0.106})\\
Extended UDS Democracy Score (Lagged) & \num{0.257} & \num{-0.085}\\
 & (\num{0.181}) & (\num{0.270})\\
Extended UDS Democracy Score\textasciicircum2 (Lagged) & \num{-0.726}* & \num{-0.761}*\\
 & (\num{0.211}) & (\num{0.352})\\
\% Mountainous Terrain (Logged) & \num{0.055} & \num{0.342}*\\
 & (\num{0.067}) & (\num{0.112})\\
Ethnic Fractionalization & \num{0.442} & \num{0.333}\\
 & (\num{0.358}) & (\num{0.554})\\
Religious Fractionalization & \num{-0.389} & \num{-0.281}\\
 & (\num{0.402}) & (\num{0.593})\\
t & \num{-0.074}+ & \num{-0.111}*\\
 & (\num{0.039}) & (\num{0.056})\\
t\textasciicircum2 & \num{0.004}* & \num{0.005}+\\
 & (\num{0.002}) & (\num{0.003})\\
t\textasciicircum3 & \num{0.000}* & \num{0.000}+\\
 & (\num{0.000}) & (\num{0.000})\\
Intercept & \num{-5.098}* & \num{-6.591}*\\
 & (\num{1.351}) & (\num{2.084})\\
\midrule
Num.Obs. & \num{8192} & \num{8192}\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05}\\
\end{tabular}
\end{table}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.

\leavevmode\hypertarget{ref-arelbundock2021m}{}%
Arel-Bundock, Vincent (2021b) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).

\leavevmode\hypertarget{ref-arelbundock2021w}{}%
Arel-Bundock, Vincent (2021a) \emph{WDI: World Development Indicators and Other World Bank Data} (\url{https://CRAN.R-project.org/package=WDI}).

>>>>

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: March 03, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
\author{true}
\date{}

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: March 03, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Steven V. Miller (Stockholm University)},
             pdfkeywords = {software, statistical analysis, peace science, reproducibility},
            pdftitle={\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
   \MakeUppercase{Steven V. Miller}, \small{Stockholm University}

@@

\texttt{\{peacesciencer\}} addresses these problems. Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers. Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch. Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. \texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts. All are done with the maximum possible transparency. The project is available for public view on Github (\url{https://github.com/svmiller/peacesciencer/}). The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created. The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

\begin{longtable}[t]{>{}lc>{\raggedright\arraybackslash}p{20em}}
\caption{\label{tab:tab-addfunctions}Functions that Add Data}\\
\toprule
\textbf{Function} & \textbf{Type} & \textbf{Description}\\
\midrule
\ttfamily{add\_archigos()} & D,S & Add Archigos political leader information to a data frame\\
\ttfamily{add\_atop\_alliance()} & D, LD & Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame\\
\ttfamily{add\_capital\_distance()} & D, L, LD, S & Add capital-to-capital distance to a data frame\\
\ttfamily{add\_ccode\_to\_gw()} & D, L, LD, S & Match CoW state codes to G-W state codes\\
\ttfamily{add\_contiguity()} & D, L, LD, S & Add CoW direct contiguity information to a data frame\\
\ttfamily{add\_cow\_alliance()} & D, LD & Add CoW alliance data to a data frame\\
\ttfamily{add\_cow\_majors()} & D, L, LD, S & Add CoW major power information to a data frame\\
\ttfamily{add\_cow\_mids()} & D & Add CoW-MID data to a dyad-year data frame\\
\ttfamily{add\_cow\_trade()} & D, L, LD, S & Add CoW trade data to a data frame\\
\ttfamily{add\_cow\_wars()} & D, S & Add CoW war data to a dyad-year or state-year data frame\\
\ttfamily{add\_creg\_fractionalization()} & D, L, LD, S & Add fractionalization/polarization estimates from CREG to a data frame\\
\ttfamily{add\_democracy()} & D, L, LD, S & Add democracy information to a data frame\\
\ttfamily{add\_fpsim()} & D, LD & Add dyadic foreign policy similarity measures to a data frame\\
\ttfamily{add\_gml\_mids()} & D, L, LD, S & Add Gibler-Miller-Little (GML) MID data to a data frame\\
\ttfamily{add\_gwcode\_to\_cow()} & D, L, LD, S & Match G-W state codes to CoW state codes\\
\ttfamily{add\_igos()} & D, S & Add CoW intergovernmental organizations (IGO) data to a data frame\\
\ttfamily{add\_lead()} & L, LD & Add leader experience/attributes (LEAD) to data\\
\ttfamily{add\_lwuf()} & L, LD & Add leader willingness to use force estimates to data\\
\ttfamily{add\_minimum\_distance()} & D, L, LD, S & Add minimum distance data to data frame\\
\ttfamily{add\_nmc()} & D, L, LD, S & Add Correlates of War National Military Capabilities data\\
\ttfamily{add\_peace\_years()} & D, S & Add 'peace years' to dyad-year/state-year conflict data\\
\ttfamily{add\_rugged\_terrain()} & D, L, LD, S & Add rugged terrain information to a data frame\\
\ttfamily{add\_sdp\_gdp()} & D, L, LD, S & Add (surplus, gross) domestic product data to a data frame\\
\ttfamily{add\_spells()} & D, L, LD, S & Add duration 'spells' to a data frame\\
\ttfamily{add\_strategic\_rivalries()} & D, S & Add strategic rivalry information to a data frame\\
\ttfamily{add\_ucdp\_acd()} & S & Add Uppsala Conflict Data Program (UCDP) Armed Conflict Data to state-year data frame\\
\ttfamily{add\_ucdp\_onsets} & S & Add UCDP onsets to state-year data\\
\ttfamily{filter\_prd()} & D, LD & Filter dyad-year data to just politically relevant dyads\\
\ttfamily{wc\_duration()} & C & Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration\\
\ttfamily{wc\_fatality()} & C & Whittle Duplicate Conflict-Years by Highest Fatality\\
\ttfamily{wc\_hostility()} & C & Whittle Duplicate Conflict-Years by Highest Hostility\\
\ttfamily{wc\_jds()} & C & Whittle Duplicate Conflict-Years by Just Dropping Something\\
\ttfamily{wc\_onsets()} & C & Whittle Uniqute Conflict Onset-Years from Conflict-Year Data\\
\ttfamily{wc\_recip()} & C & Whittle Duplicate Conflict-Years by Conflict Reciprocation\\
\ttfamily{wc\_stmon()} & C & Whittle Duplicate Conflict-Years by Lowest Start Month\\
\bottomrule
\end{longtable}
\endgroup{}

@@

\texttt{\{peacesciencer\}} is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.3. \texttt{\{peacesciencer\}} is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

\texttt{\{peacesciencer\}}'s functions work out of the box, though users should find their experience augmented by two additional downloads. First, RStudio offers an IDE that serves as a user-friendly graphical user interface (GUI) over what is, at its core, a programming language with a command-line interface. RStudio's design will make it much easier for users to experiment with \texttt{\{peacesciencer\}}'s functionality and read its documentation to assist them with the use of these functions. The second additional download is \texttt{\{tidyverse\}}, itself a suite of packages that share a common form and design (Wickham et al., 2019). \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}}, and \texttt{\{peacesciencer\}} can work without it, but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality. A user can open an R session by way of RStudio and install both packages as follows.\footnote{R users will recognize that hash signs (\texttt{\#}) allow for commenting code. Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing. In subsequent code snippets, \texttt{\#\textgreater{}} will indicate console output the user can expect from these functions.}

@@

A user who approaches \texttt{\{peacesciencer\}} with a project in mind will see they can better tailor this function to what they want. Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the Uppsala Conflict Data Program (UCDP) armed conflict data. They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward. Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).

@@

Researchers using \texttt{\{peacesciencer\}} to create data for their research project should start with one of these ``create'' functions. Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher.

@@

Researchers should see that \texttt{\{peacesciencer\}}'s functionality can scale up nicely from there. For example, the following would round out the kind of information necessary to replicate Bremer's (1992) famous ``dangerous dyads'' analysis by adding information about national material capabilities (e.g.~the composite index of national capabilities {[}CINC{]}) for both states in the dyad (\texttt{add\_nmc()}), estimates of democracy for both states in the dyad (\texttt{add\_democracy()}), information about alliance commitments in the dyad-year (\texttt{add\_cow\_alliance()}, by way of Gibler (2009)), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by Anders, Fariss \& Markowitz (2020). Whereas \texttt{add\_sdp\_gdp()} is the last command in the pipe-based workflow, the \texttt{\{peacesciencer\}} call ends by assigning to an object called \texttt{Data}. This type of assignment is done with the ``right hand'' assignment operator (i.e.~\texttt{-\textgreater{}}).

@@

If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like \texttt{haven::write\_dta(Data,\ "my-data.dta")} and import it into Stata when they are done. No matter, \texttt{\{peacesciencer\}} has pre-processed, cleaned, recoded, and merged the desired data that have greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding \texttt{-9s} in these data to be \texttt{NA} in the capabilities data. In this particular application, it has already created the main data required for a replication of Bremer (1992). There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some ``weak-link'' indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the Marquez (2016) method for extending the Unified Democracy Scores {[}UDS{]} data), and the GDP per capita in the dyad. Table \ref{tab:tab-dd} is a formatted version of the results of a logistic regression model of conflict onset using these ``dangerous dyads'' indicators and temporal adjustment variables (\emph{t}). Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but \texttt{\{peacesciencer\}} greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

@@

\texttt{\{peacesciencer\}} is capable of generating data for replications of analyses at multiple levels beyond dyad-year. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{},\ subset\_years\ =\ c(1946:2019))}, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in \texttt{\{peacesciencer\}} include just the observations in that time frame). Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states. \texttt{add\_ucdp\_acd()} takes three arguments: \texttt{type}, \texttt{issue}, and \texttt{only\_wars}. \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information. Options include ``extrasystemic'', ``interstate'', ``intrastate'', and ``II'' (short for ``internationalized intrastate''). If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}). \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include ``territory'', ``government'', and ``both''. If no \texttt{issue} is specified, the function returns information for all conflicts regardless of the particular issue. \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}. The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE)} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}

@@

Finally, we can add some covariates of interest to these data. \texttt{add\_spells()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}. \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) UDS data). \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. \texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss \& Markowitz (2020) simulations. Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by Nunn \& Puga (2012) and the second is the Gibler \& Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon \& Laitin (2003)). At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.

@@

\caption{\label{tab:tab-cw}A Civil Conflict Analysis of Gleditsch-Ward State-Years in \{peacesciencer\}}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8cm}cc}
\toprule
\textbf{ } & \textbf{All UCDP Conflicts} & \textbf{Wars Only}\\
\midrule
GDP per Capita (Lagged) & \num{-0.285}* & \num{-0.343}*\\
 & (\num{0.110}) & (\num{0.172})\\
Population Size (Lagged) & \num{0.229}* & \num{0.272}*\\
 & (\num{0.067}) & (\num{0.106})\\
Extended UDS (Lagged) & \num{0.257} & \num{-0.085}\\
 & (\num{0.181}) & (\num{0.270})\\
Extended UDS\textasciicircum2 (Lagged) & \num{-0.726}* & \num{-0.761}*\\
 & (\num{0.211}) & (\num{0.352})\\
\% Mountainous Terrain (Logged) & \num{0.055} & \num{0.342}*\\
 & (\num{0.067}) & (\num{0.112})\\
Ethnic Fractionalization & \num{0.442} & \num{0.333}\\
 & (\num{0.358}) & (\num{0.554})\\
Religious Fractionalization & \num{-0.389} & \num{-0.281}\\
 & (\num{0.402}) & (\num{0.593})\\
t & \num{-0.074}+ & \num{-0.111}*\\
 & (\num{0.039}) & (\num{0.056})\\
t\textasciicircum2 & \num{0.004}* & \num{0.005}+\\
 & (\num{0.002}) & (\num{0.003})\\
t\textasciicircum3 & \num{0.000}* & \num{0.000}+\\
 & (\num{0.000}) & (\num{0.000})\\
Intercept & \num{-5.098}* & \num{-6.591}*\\
 & (\num{1.351}) & (\num{2.084})\\
\midrule
Num.Obs. & \num{8192} & \num{8192}\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05}\\
\end{tabular}
\end{table}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-andersetal2020bbgb}{}}%
Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.

\leavevmode\vadjust pre{\hypertarget{ref-arelbundock2021m}{}}%
Arel-Bundock, Vincent (2021b) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).

\leavevmode\vadjust pre{\hypertarget{ref-arelbundock2021w}{}}%
Arel-Bundock, Vincent (2021a) \emph{WDI: World Development Indicators and Other World Bank Data} (\url{https://CRAN.R-project.org/package=WDI}).
